Transcriptomic Analysis for Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-07-03 Cover_letter_JPM.tex
Updates from Overleaf

The manuscript posted in the preprint server, ResearchSqaure, is our first version (2020/09/28) of current manuscript. 
Since previous reviewer's suggestion, we made a major revision on the followings:
1) suggested candidates of biomarker has been modified  by additional validation using an independent patient cohort (GSE2837).
We also reviewed literature that supports the prognostic impact of these candidate genes. The resulting Figure 4, Table 3 and Table 4 is different between the two manuscripts.
2) Since that revision, we added holistic cancer care in the discussions section (please see Answer 1-2).% \ref{page:1-2}).

@@

% problem of it (少 n=30, and RFS only: validation rate 14/100 而且 10% cutoff)
%x Prognoscan (GSE2837) (n=30 less)  RFS only:  10\% cutoff to yield a \textit{P}-value < 0.05.\\
%找出  3 candidates
During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.
%
% [2021/06/13-17]
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort, CAMK2N1, CALML5, and FCGBP, keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox'HR ($>1.8$ or $<0.65$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002 
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the "top rank" of the other 17 genes is replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
There is consensus between the TCGA and GSE65858 cohorts that CAMK2N1, CALML5, and FCGBP are significant candidates for the biomarker.
%In the consensus of TCGA and GSE65858 cohorts,

@@

% [2021/06/13-17] GSE65858
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort with median cutoffs, CAMK2N1, CALML5, and FCGBP (3 of the 20 genes), keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox's HR ($>1.8$ or $<0.6$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the significance of the other 17 genes are replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
%This GSE65858 cohort has been applied for selection to generate our candidate genes: \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19} (two probes), and , \textcolor{red}{FCGBP}.
Conversely, there are 22 genes, which have greater hazard ratio ($> 1.8$ or $<0.6$) in GSE65858 cohort (Figure \ref{fig:hazards534}), drop their hazard ratio between 0.6 and 1.5 in the study of TCGA HNSCC (Figure \ref{fig:hazards3}).
Thus, there is a consensus between the TCGA and GSE65858 cohorts that CAMK2N1, CALML5, and FCGBP are significant candidates for the HNSCC biomarker.
Finally, these 3 candidates were also confirmed by a HNSCC study using the GSE2837 dataset. Please see the Kaplan-Meier plots in Supplementary Figure S2.%\ref{fig:fig_GSE2837}.
"\\[0.3cm]
%The consensus between the TCGA and GSE65858 cohorts in Kaplan-Meier survival and Cox's model
%by GEO GSE2837 HNSCC dataset (PrognoScan)
%After the discovery by our workflow, we validated the those 20 genes in the other independent \acrshort{hnscc} cohort (GSE2837).
%The survival significance (Kaplan-Meier \textit{P} value) is in the following probes among 10 genes:\\
%1) poor prognosis: CAMK2N1 (0.048214), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587);\\ % FOXA2 (0.038125)
%2) better prognosis: IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);\\
%17 out of 20 candidates

@@

%2) \\
%(ok) % pvalueTex is essential; find smaller P value to predict large effect size (HR)
% The effect size is the main finding of a quantitative study.
There is a benefit for using sliding-window cutoff method (between $30\% - 70\%$ quantile) in Kaplan-Meier analysis of TCGA HNSCC cohort at the beginning.
We compared the results of cutoff at optimal \textit{P} value by sliding window or just at the median of gene expression.
It shows that the number of genes with unadjusted \textit{P} value $< 0.05$, which is 6284 versus 3118, respectively. After FDR correction, it becomes 967 versus 209, respectively.
% HNSCC_OS_marginS_THREE_pvalue005_noCancerGene
%Result by optimal cutoff: total 9416 plots,
%6284 are raw KM P value < 0.05
%534?? genes are FDR KM P value <0.05
%967 genes which HR is greater than 1.5 or less than 0.5.
%Result by median cutoff: total 20080 genes, 
%3118 are raw KM P value < 0.05; 
%209 genes are FDR KM P value <0.05
The sliding-window cutoff method could catch more potential candidates, which have \textit{P} values far less than $0.001$, for subsequent Cox's modeling.
That is because of the properly selected cutoff improving the statistical significance.
To find a smaller \textit{P} value might predict large effect size (HR) associated biomarkers.
Then, these preliminary candidates will have an opportunity to be carefully selected by using Bonferroni correction, their effect size (Cox's HR), and the other independent cohorts (GSE65858, GSE2837) to prevent the false discovery.

@@

We explain aforementioned situation by examples.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary figure S1, a "S"- or "W"-shaped \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858 and HR = 1.42 in GSE2837) less than 1.8.
%%%

@@

\begin{figure}[ht]
%    \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=6cm]{rplot_KMplot_HK1.pdf}
    \caption{Kaplan-Meier plot}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width=8cm]{Rplot_TCGA_HNSCC_CoxHR_CAMK2N1_signature4_FDRKM.pdf}  
    \caption{Volcano plot}
  \end{subfigure}
%However, # 4 gene signature: #### 
%# "EZR" has \textit{P} value $>0.05$
  \caption{(a) OS in TCGA HNSCC for a gene signature including HK1, EZR, RRAGA, and DUSP16 (Kaplan-Meier \textit{P} value = \num{1.31e-6}); (b) Those genes (listed on the rigth side) have not significant hazard ratio (HR) if calculated individually (EZR is not shown on volcano plot due to it's Kaplan-Meier \textit{P} value $>0.05$)}
    \label{fig:signature4}
\end{figure}

>>>>

The manuscript posted in the preprint server, ResearchSqaure, is our first version (2020/09/28) of current manuscript. 
Since previous reviewer's suggestion, we made a major revision on the followings:
1) suggested candidates of biomarker has been modified  by additional validation using an independent patient cohort (GSE65858).
%We also reviewed literature that supports the prognostic impact of these candidate genes. 
The resulting Figure 4, and Table 3 are different between the two manuscripts.
2) Since that revision, we added holistic cancer care in the discussions section (please see Answer 1-2).% \ref{page:1-2}).

@@

% problem of it (少 n=30, and RFS only: validation rate 14/100 而且 10% cutoff)
%x Prognoscan (GSE2837) (n=30 less)  RFS only:  10\% cutoff to yield a \textit{P}-value < 0.05.\\
%找出  3 candidates
During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.
%
% [2021/06/13-17]
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort, CAMK2N1, CALML5, and FCGBP, keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox'HR ($>1.8$ or $<0.6$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002 
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the "top rank" of the other 17 genes is replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
There is consensus between the TCGA and GSE65858 cohorts that CAMK2N1, CALML5, and FCGBP are significant candidates for the biomarker.
%In the consensus of TCGA and GSE65858 cohorts,

@@

% [2021/06/13-17] GSE65858
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort with median cutoffs, CAMK2N1, CALML5, and FCGBP (3 of the 20 genes), keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox's HR ($>1.8$ or $<0.6$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the significance of the other 17 genes are replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
%This GSE65858 cohort has been applied for selection to generate our candidate genes: \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19} (two probes), and , \textcolor{red}{FCGBP}.
Conversely, there are 22 genes, which have greater hazard ratio ($> 1.8$ or $<0.6$) in GSE65858 cohort (Figure \ref{fig:hazards534}), drop their hazard ratio between 0.6 and 1.5 in the study of TCGA HNSCC (Figure \ref{fig:hazards3}).
Thus, there is a consensus between the TCGA and GSE65858 cohorts that CAMK2N1, CALML5, and FCGBP are significant candidates for the HNSCC biomarker.
%Finally, these 3 candidates were also confirmed by a HNSCC study using the GSE2837 dataset. Please see the Kaplan-Meier plots in Supplementary Figure S2.%\ref{fig:fig_GSE2837}.
"\\[0.3cm]
%The consensus between the TCGA and GSE65858 cohorts in Kaplan-Meier survival and Cox's model
%by GEO GSE2837 HNSCC dataset (PrognoScan)
%After the discovery by our workflow, we validated the those 20 genes in the other independent \acrshort{hnscc} cohort (GSE2837).
%The survival significance (Kaplan-Meier \textit{P} value) is in the following probes among 10 genes:\\
%1) poor prognosis: CAMK2N1 (0.048214), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587);\\ % FOXA2 (0.038125)
%2) better prognosis: IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);\\
%17 out of 20 candidates

@@

%2) \\
%(ok) % pvalueTex is essential; find smaller P value to predict large effect size (HR)
% The effect size is the main finding of a quantitative study.
There is a benefit for using sliding-window cutoff method (between $30\% - 70\%$ quantile) in Kaplan-Meier analysis of TCGA HNSCC cohort at the beginning.
We compared the results of cutoff at optimal \textit{P} value by sliding window or just at the median of gene expression.
It shows that the number of genes with unadjusted \textit{P} value $< 0.05$, which is 6284 versus 3118, respectively. After FDR correction, it becomes 967 versus 209, respectively.
% HNSCC_OS_marginS_THREE_pvalue005_noCancerGene
%Result by optimal cutoff: total 9416 plots,
%6284 are raw KM P value < 0.05
%534?? genes are FDR KM P value <0.05
%967 genes which HR is greater than 1.5 or less than 0.5.
%Result by median cutoff: total 20080 genes, 
%3118 are raw KM P value < 0.05; 
%209 genes are FDR KM P value <0.05
The sliding-window cutoff method could catch more potential candidates, which have \textit{P} values far less than $0.001$, for subsequent Cox's modeling.
That is because of the properly selected cutoff improving the statistical significance.
To find a smaller \textit{P} value might predict large effect size (HR) associated biomarkers.
Then, these preliminary candidates will have an opportunity to be carefully selected by using Bonferroni correction, their effect size (Cox's HR), and the other independent cohort (GSE65858) to prevent the false discovery.

@@

We explain aforementioned situation by examples.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary figure S1, a "S"- or "W"-shaped \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858) less than 1.8.
%%%

@@

\begin{figure}[ht]
%    \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=6cm]{rplot_KMplot_HK1.pdf}
    \caption{Kaplan-Meier plot}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width=8cm]{Rplot_TCGA_HNSCC_CoxHR_CAMK2N1_signature4_FDRKM.pdf}  
    \caption{Volcano plot}
  \end{subfigure}
%However, # 4 gene signature: #### 
%# "EZR" has \textit{P} value $>0.05$
  \caption{(a) Overall survival in TCGA HNSCC for a gene signature including HK1, EZR, RRAGA, and DUSP16 (Kaplan-Meier \textit{P} value = \num{1.31e-6}); (b) Those genes (listed on the right side) have not significant hazard ratio (HR) if calculated individually (EZR is not shown on volcano plot due to it's Kaplan-Meier \textit{P} value $>0.05$)}
    \label{fig:signature4}
\end{figure}

### 2021-07-03 Supplementary_figures.tex
Updates from Overleaf

\section{STables}
\underline{Supplementary figure legends}\\[1cm]
%ST1
Supplementary Table \ref{table:table1}\\
GSE2837 query results from PrognoScan: Kaplan-Meier plots of 10 genes (the cutoff of high risk and low risk groups, which is derived from cumulative P-value plots). The poor prognostic genes are marked as red; the better prognostic genes are marked as green.

@@

\multicolumn{9}{|l!{\color{black}\vrule}}{Remark: \checkmark validated by GSE2837}                                                                                                                                                                                                                                                                                                                                                             \\

@@

\multicolumn{9}{|l|}{\begin{tabular}[c]{@{}l@{}}Selection criteria:~\\~Kaplan-Meier Bonferroni-adjusted \textit{P} value \textless{} 0.05~\\~Cox's univariate and multivariate HR \textgreater{}= 1.5\\

@@

lncRNA: Long non-coding RNA\\Remark: \checkmark Validated by GSE2837 ~\end{tabular}}                                                                                                                                                                                                                 \\

@@

%S2
Supplementary Figure \ref{fig:fig_GSE2837}\\
GSE2837 query results from PrognoScan: Kaplan-Meier plots of 10 genes (the cutoff of high risk and low risk groups, which is derived from cumulative P-value plots). The poor prognostic genes are marked as red; the better prognostic genes are marked as green.

@@

% S2
% from Answer2-2 GSE2837; 
\begin{figure}
\raggedleft
\includegraphics[width=12cm]{Answer_2-2.pdf}
\caption{GSE2837 query results from PrognoScan: Kaplan-Meier plots of 10 genes (the cutoff of \textcolor{red}{high risk} and \textcolor{blue}{low risk} groups, which is derived from cumulative P-value plots). The poor prognostic genes are marked as \textcolor{red}{red}; the better prognostic genes are marked as \textcolor{green}{green}.}
\label{fig:fig_GSE2837}
\end{figure}

>>>>

\section{STables}
\underline{Supplementary figure legends}\\[1cm]
%ST1
%Supplementary Table \ref{table:table1}\\
%GSE2837 query results from PrognoScan: Kaplan-Meier plots of 10 genes (the cutoff of high risk and low risk groups, which is derived from cumulative P-value plots). The poor prognostic genes are marked as red; the better prognostic genes are marked as green.

@@

\multicolumn{9}{|l!{\color{black}\vrule}}{}%Remark: \checkmark validated by GSE2837}                                                                                                                                                                                                                                                                                                                                                             \\

@@

\multicolumn{9}{|l|}{\begin{tabular}[c]{@{}l@{}}Selection criteria:~\\~Kaplan-Meier Bonferroni-adjusted \textit{P} value \textless{} 0.05~\\~Cox's univariate and multivariate HR \textless{}= 1.5\\

@@

lncRNA: Long non-coding RNA\\%Remark: \checkmark Validated by GSE2837
~\end{tabular}}                                                                                                                                                                                                                 \\

@@

%S2
%Supplementary Figure \ref{fig:fig_GSE2837}\\
%GSE2837 query results from PrognoScan: Kaplan-Meier plots of 10 genes (the cutoff of high risk and low risk groups, which is derived from cumulative P-value plots). The poor prognostic genes are marked as red; the better prognostic genes are marked as green.

@@

% S2
% from Answer2-2 GSE2837; 
%\begin{figure}
%\raggedleft
%\includegraphics[width=12cm]{Answer_2-2.pdf}
%\caption{GSE2837 query results from PrognoScan: Kaplan-Meier plots of 10 genes (the cutoff of \textcolor{red}{high risk} and \textcolor{blue}{low risk} groups, which is derived from cumulative P-value plots). The poor prognostic genes are marked as \textcolor{red}{red}; the better prognostic genes are marked as \textcolor{green}{green}.}
%\label{fig:fig_GSE2837}
%\end{figure}
